---
title: "Peer assessment 1 (Reproducible Research)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hold')
```

## Loading and preprocessing the data

```{r loaddata}
library(ggplot2)
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}

totalstep <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
qplot(totalstep, binwidth=1000, xlab="steps taken each day")
mean(totalstep, na.rm=TRUE)
median(totalstep, na.rm=TRUE)
```

## Average daily activity pattern

```{r}
averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken")
```

On average across all the days in the dataset, the 5-minute interval contains
the maximum number of steps?

```{r}
averages[which.max(averages$steps),]
```

## Imputing missing values

```{r how_many_missing}
missing <- is.na(data$steps)
table(missing)
```

Fill the missing values with the mean valu.

```{r}
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```

Histogram of the total number of steps taken each day and calculate the mean and median total number of steps.

```{r}
totalstep <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(totalstep, binwidth=1000, xlab="total number of steps taken each day")
mean(totalstep)
median(totalstep)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("lunes", "martes", "miércoles", "jueves", "viernes")) 
        return("weekday") else if (day %in% c("sábado", "domingo")) 
        return("weekend") else stop("invalid date")
}

filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN = weekday.or.weekend)

averages <- aggregate(steps ~ interval + day, data = filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + 
    xlab("5-minute interval") + ylab("Number of steps")
```