---
title: "Peer assessment 1 (Reproducible Research)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hold')
```

## Loading and preprocessing the data

Show any code that is needed to Load the data (i.e. read.csv())
Process/transform the data (if necessary) into a format suitable for your analysis

```{r}
library(ggplot2)
library(dplyr)
setwd("~/R/RepData_PeerAssessment1-master")
list.files()
activity <- read.csv("activity.csv", stringsAsFactors = F)
str(activity)
```

## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day.
Make a histogram of the total number of steps taken each day.
Mean and median of the total number of steps taken per day.

```{r}
total.steps <- tapply(activity$steps, activity$date, FUN = sum, na.rm = TRUE)
qplot(total.steps, binwidth = 1000, xlab = "total number of steps taken each day")

step_mean <- mean(total.steps, na.rm = TRUE)
step_median <- median(total.steps, na.rm = TRUE)
step_mean
step_median
``` 

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis). Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
  averages <- aggregate(x=list(steps=activity$steps),
                        by=list(interval=activity$interval), 
                        FUN=mean, na.rm = TRUE)
  ggplot(data = averages, aes(x = interval, y = steps)) + geom_line() + xlab("5_minute_interval") + 
      ylab("average number of steps taken")
  
averages[which.max(averages$steps), ]
```

##Imputing missing values 

Devise a strategy for filling in all of the missing values in the dataset. 
Create a new dataset that is equal to the original dataset but with the missing data filled in.
Make a histogram 

```{r}
str(activity)
fill <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps)) 
        filled <- c(steps) else filled <- (averages[averages$interval == interval, "steps"])
    return(filled)
}
filled.data <- activity
filled.data$steps <- mapply(fill, filled.data$steps, filled.data$interval)

total.steps <- tapply(filled.data$steps, filled.data$date, FUN = sum)
qplot(total.steps, binwidth = 1000, xlab = "total number of steps taken each day")

mean(total.steps)
median(total.steps)

```

## Differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.
In this part I create new variables weekday and DayType to find if weekend or wekday.
The I do the plots. 


```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("lunes", "martes", "miércoles", "jueves", "viernes")) 
        return("weekday") else if (day %in% c("sábado", "domingo")) 
        return("weekend") else stop("invalid date")
}

filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN = weekday.or.weekend)

averages <- aggregate(steps ~ interval + day, data = filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + 
    xlab("5-minute interval") + ylab("Number of steps")

head(filled.data)
table(filled.data$day)

```